Uncertainty on the galaxy-halo connection for Lyman-$\alpha$ emitters at $z=3.1$

### 2016-03-20 paper/ms.tex
latex setup update

\documentclass{emulateapj} 
%\documentclass[a4,useAMS,usenatbib,usegraphicx]{mn2e}

%External Packages and personalized macros
%=========================================================================
%		EXTERNAL PACKAGES
%=========================================================================
%\usepackage{amsmath} 
%\usepackage{amssymb} 
%\usepackage[section]{placeins}

\usepackage {graphicx}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{float}
\bibliographystyle{apj}

\usepackage {graphicx}

%\usepackage{graphics}
%\usepackage[dvips]{epsfig}
%\usepackage{epsfig}  
\usepackage{color}
\usepackage[normalem]{ulem}
\usepackage{hyperref}
\usepackage{caption}
%Non reposionated tables
%\usepackage{float}
\restylefloat{table}
%Multiple columns support for tables
\usepackage{array}
\usepackage{booktabs}
\setlength{\heavyrulewidth}{1.5pt}
\setlength{\abovetopsep}{4pt}
\pdfminorversion=5

@@

\newcommand{\LCDM}{$\Lambda$CDM~}
\newcommand{\beq}{\begin{eqnarray}}  
\newcommand{\eeq}{\end{eqnarray}}  
\newcommand{\zz}{$z\sim 3$} 
%\newcommand{\apj}{ApJ}  
%\newcommand{\apjs}{ApJS}  
%\newcommand{\apjl}{ApJL}  
%\newcommand{\aj}{AJ}  
%\newcommand{\mnras}{MNRAS}  
%\newcommand{\mnrassub}{MNRAS accepted}  
%\newcommand{\aap}{A\&A}  
%\newcommand{\aaps}{A\&AS}  
%\newcommand{\araa}{ARA\&A}  
%\newcommand{\nat}{Nature}  
%\newcommand{\physrep}{PhR}
%\newcommand{\pasp}{PASP}    
%\newcommand{\pasj}{PASJ}    
\newcommand{\avg}[1]{\langle{#1}\rangle}  
\newcommand{\ly}{{\ifmmode{{\rm Ly}\alpha}\else{Ly$\alpha$}\fi}}
\newcommand{\hMpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc}\fi}}  
\newcommand{\hGpc}{{\ifmmode{h^{-1}{\rm Gpc}}\else{$h^{-1}$Gpc}\fi}}  
\newcommand{\hmpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc}\fi}}  
\newcommand{\hkpc}{{\ifmmode{h^{-1}{\rm kpc}}\else{$h^{-1}$kpc}\fi}}  
\newcommand{\hMsun}{{\ifmmode{h^{-1}{\rm {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}  
\newcommand{\Mmin}{{\ifmmode{{M_{\rm min}}}\else{${M_{\rm min}}$}\fi}}
\newcommand{\mmin}{{\ifmmode{{M_{\rm min}}}\else{${M_{\rm min}}$}\fi}}
\newcommand{\mmax}{{\ifmmode{{M_{\rm max}}}\else{${M_{\rm max}}$}\fi}}
\newcommand{\lmmin}{{\ifmmode{{\log M_{\rm min}}}\else{${\log M_{\rm min}}$}\fi}}
\newcommand{\lmmax}{{\ifmmode{{\log M_{\rm max}}}\else{${\log M_{\rm max}}$}\fi}}

@@

\newcommand{\Msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\lya}{{Lyman$\alpha$~}}
\newcommand{\clara}{{\texttt{CLARA}}~}
\newcommand{\rand}{{\ifmmode{{\mathcal{R}}}\else{${\mathcal{R}}$ }\fi}}  
%SAMPLES
\newcommand{\GHBDM}{\texttt{GH}$_{\mbox{\tiny{BDM}}}$ }
\newcommand{\GHFOF}{\texttt{GH}$_{\mbox{\tiny{FOF}}}$ }
\newcommand{\IHBDM}{\texttt{IH}$_{\mbox{\tiny{BDM}}}$ }
\newcommand{\IHFOF}{\texttt{IH}$_{\mbox{\tiny{FOF}}}$ }
\newcommand{\PBDM}{\texttt{P}$_{\mbox{\tiny{BDM}}}$ }
\newcommand{\PFOF}{\texttt{P}$_{\mbox{\tiny{FOF}}}$ }
\newcommand{\IPBDM}{\texttt{IP}$_{\mbox{\tiny{BDM}}}$ }
\newcommand{\IPFOF}{\texttt{IP}$_{\mbox{\tiny{FOF}}}$ }
\newcommand{\RIPBDM}{\texttt{RIP}$_{\mbox{\tiny{BDM}}}$ }
\newcommand{\RIPFOF}{\texttt{RIP}$_{\mbox{\tiny{FOF}}}$ }

@@

\begin{document}

%=========================================================================
%		FRONT MATTER
%=========================================================================
\title{Uncertainty on the galaxy-halo connection for Lyman-$\alpha$ emitters at $z=3.1$}
\author{
  Juli\'an E. Mej\'ia-Restrepo \thanks{jemejia@das.uchile.cl}$^{1,3}$,
  Jaime E. Forero-Romero \thanks{je.forero@uniandes.edu.co}$^{2}$ 
}

@@

\begin{abstract}
We study the impact of cosmic variance and observational uncertainties in constraining the mass and  occupation fraction ($f_{\rm occ}$) of dark matter halos hosting \ly\ Emitting  Galaxies at high redshift (LAEs).
We use a N-body simulation to construct mock catalogs with the same typical size  of observed fields at $z=3.1$ ($\sim 1 {\rm deg^2}$) to match the observed  angular correlation function (ACF) and number density of LAEs.  
In our model a dark matter halo with mass in the range $M_{\rm min}<M_{\rm h}<M_{\rm   max}$ can only host one detectable LAE at most. By following a thorough Markov Chain Monte-Carlo exploration of the parameter space determined by $M_{\rm min}$ and $M_{\rm   max}$, our analysis shows that $f_{\rm occ}$ is uniquely determined by $M_{\rm min}$ regardless of $M_{\rm max}$ using the relation $\focc\sim0.1\left(\mmin/10^{10.5}\right)^{0.93}$ . 
However, the current observational data only allows us to put weak constraints on $M_{\rm min}$, $M_{\rm max}$ and consequently on $f_{\rm occ}$. 
Particularly, we find that $10^{9.6}\hMsun\leq M_{\rm min}\leq 10^{11.0}\hMsun$ , $10^{10.9}\hMsun\leq M_{\rm max}\leq 10^{13.0}\hMsun$ and $0.02\leq f_{\rm occ}\leq 0.30$. 
Nevertheless, we also find that  the upcoming large surveys can help to put tighter constraints on $M_{\rm min}$ and $f_{\rm occ}$ through the width measurement of the LAE number distribution function obtained over several fields of the same size of current observations. 
Analogously, the  improvement in the precision of the measured ACF will assist in constraining $M_{\rm max}$.  
\end{abstract}

@@

Then we say that a halo will host a LAE with probability $f_{\rm occ}$ if and only if the halo mass is in the range $M_{\rm min}< M_{\rm h} < M_{\rm max}$.
The probability $f_{\rm occ}$ can be thought as the occupation fraction of halos which can be automatically set as the ration of the observed number of LAEs to the number of halos within the considered mass range, that is $f_{\rm occ}\equiv N_{\rm LAE}/N_{\rm halos}$.

@@

We use the angular correlation function to compare the observations with our mock catalogs. 
This is done trough a thorough exploration of the parameter space of the models ${\mathcal M}$  by  means of a Monte Carlo Markov Chain minimization using  the EMCEE python package \citep{emcee2013}.
We put a flat prior on $\log M_{\rm min}$ and $\log \mmax$ to vary between  $9.2$ up to $13.4$, corresponding to the halo mass range of the simulation at $z=3$. 
Given that the typical scatter  in $N_{\rm halos}$ is about 0.3 dex (i.e. about  a factor of 2), our parameter space is restricted to models where the median number of dark matter halos is larger than  $N_{\rm LAE}/3$. 
The MCMC exploration is done using a total of 24 seeds and 400 iterations (9600 models) to sample the posterior PDF,  $P(\mmin, \mmax, f_{\rm occ}| \mathcal{M}) \propto \exp(-\chi_{\mathcal M}^2/2)$, with

@@

In Fig. \ref{fig:like} we present yhe one and two dimensional projections of the posterior probability distributions of \mmin, \mmax and \focc. We find that the \mmin and \mmax parameter space cannot be tightly constrained mainly because of cosmic variance and the large observational Poisonnian uncertainty in the ACF. We however found that 9.6$<\log \mmin<$11.0 and 10.9$<\log\mmax<$13.0.   We interestingly find that \focc\ is basically completely determined by \mmin  going from \focc$=$0.015 when $\log\mmin=$9.6 to \focc$=$0.30 when $\log\mmin=$11.0  ($\focc\sim0.1\left(\mmin/10^{10.5}\right)^{0.93}$). We remark that those models where $\log \focc>0.00$ ($\focc>1$) correspond 
to models where the number density  of halos is smaller than the number of observed LAEs but are
still consistent because of the uncertainty in the median number density of LAEs in the universe due to cosmic variance (see Fig. \ref{fig:cosmicv0}).%in the number of dark matter halos along the mock fields

@@

 In Fig. \ref{fig:mmed} the green vertical lines represent  the 16, 50 and 84 percentiles  of the posterior probability distribution of \mmin that we previously obtained. From this figure we find that the median dark matter halo mass is $\tilde{M}_{\rm h} = 10^{10.6^{+0.7}_{-0.6}$. This result is consistent with previous estimations of  the  median dark matter halo mass reported by  \citet{Bielby16} ($\tilde{M}_{\rm h} = 10^{11.0\pm0.6}$), \citet{Gawiser07} ($\tilde{M}_{\rm h} = 10^{10.9\pm0.9}$) and \citet{Ouchi2010} ($\tilde{M}_{\rm h} = 6.7^{+42.0}_{6.7}\times10^{10}}$) using semi-analitical approaches.

@@

\begin{figure}
\includegraphics[width=0.46\textwidth, height=0.33\textwidth]{mmin_dfocc1.png}
\caption{Width of the halo number distribution over the 54 mock field of the simulation ($\Delta \log \left( N_{\rm halos}/\tilde{N}_{\rm halos}\right)$) of the central 68 (blue diamonds) and  95(red circles)  percentiles vs $M_{\rm min}$. Points are color coded according to their associated value of $\Delta M \equiv M_{\rm min}-M_{\rm min}$. The green line diagonal line accounts for the 1:1 relation and the green vertical lines represent  the 16, 50 and 84 percentiles  of the posterior probability distribution of \mmin.}
\label{fig:cosmicv}
\end{figure}

@@

In this letter we studied the impact of cosmic variance and observational uncertainties in constraining the mass range and  occupation fraction of dark matter halos hosting  LAEs.  
To this end, we used the BolshoiP N-body simulation to construct  27 mock fields  with the same typical size  of observed fields at  $z=3.1$ ($\sim 1 {\rm deg^2}$).  
In our model a dark matter halo with mass in the range $M_{\rm min}<M_{\rm h}<M_{\rm   max}$ can only host one detectable LAE at most. 
We explored the parameter space determined by \mmin\ and \mmax\ using Monte Carlo Markov-Chain minimization to match the observed  ACF and mean number density of LAEs.

>>>>

\documentclass{emulateapj}
\submitted{{\it Submitted for publication in ApJ}}
\usepackage{multirow,color,wrapfig,ulem}
\usepackage {graphicx}
\usepackage{graphics}
\usepackage[dvips]{epsfig}

@@

\newcommand{\LCDM}{$\Lambda$CDM~}
\newcommand{\beq}{\begin{eqnarray}}  
\newcommand{\eeq}{\end{eqnarray}}  
\newcommand{\zz}{$z\sim 3$} 
\newcommand{\avg}[1]{\langle{#1}\rangle}  
\newcommand{\ly}{{\ifmmode{{\rm Ly}\alpha}\else{Ly$\alpha$}\fi}}
\newcommand{\hMpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc}\fi}}  
\newcommand{\hGpc}{{\ifmmode{h^{-1}{\rm Gpc}}\else{$h^{-1}$Gpc}\fi}}  
\newcommand{\hmpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc}\fi}}  
\newcommand{\hkpc}{{\ifmmode{h^{-1}{\rm kpc}}\else{$h^{-1}$kpc}\fi}}  
\newcommand{\hMsun}{{\ifmmode{h^{-1}{\rm {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}  
\newcommand{\Mmin}{{\ifmmode{{M_{\rm min}}}\else{${M_{\rm min}}$}\fi}}
\newcommand{\mmin}{{\ifmmode{{M_{\rm min}}}\else{${M_{\rm min}}$}\fi}}
\newcommand{\mmax}{{\ifmmode{{M_{\rm max}}}\else{${M_{\rm max}}$}\fi}}
\newcommand{\lmmin}{{\ifmmode{{\log M_{\rm min}}}\else{${\log M_{\rm min}}$}\fi}}
\newcommand{\lmmax}{{\ifmmode{{\log M_{\rm max}}}\else{${\log M_{\rm max}}$}\fi}}

@@

\newcommand{\Msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\lya}{{Lyman$\alpha$~}}
\newcommand{\clara}{{\texttt{CLARA}}~}
\newcommand{\rand}{{\ifmmode{{\mathcal{R}}}\else{${\mathcal{R}}$ }\fi}}  
%SAMPLES

@@

\begin{document}
%=========================================================================
%		FRONT MATTER
%=========================================================================
\title{Uncertainty on the galaxy-halo connection for Lyman-$\alpha$ emitters at $z=3.1$}
\author{
  Juli\'an E. Mej\'ia-Restrepo \thanks{jemejia@das.uchile.cl}$^{1,3}$,
  Jaime E. Forero-Romero \thanks{je.forero@uniandes.edu.co}$^{2}$ 
}

@@

\begin{abstract}
We study the impact of cosmic variance and observational uncertainties in constraining the mass and  occupation fraction ($f_{\rm occ}$) of dark matter halos hosting \ly\ Emitting  Galaxies at high redshift (LAEs).
We use a N-body simulation to construct mock catalogs with the same typical size  of observed fields at $z=3.1$ ($\sim 1 {\rm deg^2}$) to match the observed  angular correlation function (ACF) and number density of LAEs.  
In our model a dark matter halo with mass in the range $M_{\rm min}<M_{\mathrm h}<M_{\rm   max}$ can only host one detectable LAE at most. By following a thorough Markov Chain Monte-Carlo exploration of the parameter space determined by $M_{\rm min}$ and $M_{\rm   max}$, our analysis shows that $f_{\rm occ}$ is uniquely determined by $M_{\rm min}$ regardless of $M_{\rm max}$ using the relation $\focc\sim0.1\left(\mmin/10^{10.5}\right)^{0.93}$ . 
However, the current observational data only allows us to put weak constraints on $M_{\rm min}$, $M_{\rm max}$ and consequently on $f_{\rm occ}$. 
Particularly, we find that $10^{9.6}\hMsun\leq M_{\rm min}\leq 10^{11.0}\hMsun$ , $10^{10.9}\hMsun\leq M_{\rm max}\leq 10^{13.0}\hMsun$ and $0.02\leq f_{\rm occ}\leq 0.30$. 
Nevertheless, we also find that  the upcoming large surveys can help to put tighter constraints on $M_{\rm min}$ and $f_{\rm occ}$ through the width measurement of the LAE number distribution function obtained over several fields of the same size of current observations. 
Analogously, the  improvement in the precision of the measured ACF will assist in constraining $M_{\rm max}$.  
\end{abstract}

@@

Then we say that a halo will host a LAE with probability $f_{\rm occ}$ if and only if the halo mass is in the range $M_{\rm min}< M_{\mathrm h} < M_{\rm max}$.
The probability $f_{\rm occ}$ can be thought as the occupation fraction of halos which can be automatically set as the ration of the observed number of LAEs to the number of halos within the considered mass range, that is $f_{\rm occ}\equiv N_{\rm LAE}/N_{\mathrm halos}$.

@@

We use the angular correlation function to compare the observations with our mock catalogs. 
This is done trough a thorough exploration of the parameter space of the models ${\mathcal M}$  by  means of a Monte Carlo Markov Chain minimization using  the EMCEE python package \citep{emcee2013}.
We put a flat prior on $\log M_{\rm min}$ and $\log \mmax$ to vary between  $9.2$ up to $13.4$, corresponding to the halo mass range of the simulation at $z=3$. 
Given that the typical scatter  in $N_{\mathrm halos}$ is about 0.3 dex (i.e. about  a factor of 2), our parameter space is restricted to models where the median number of dark matter halos is larger than  $N_{\rm LAE}/3$. 
The MCMC exploration is done using a total of 24 seeds and 400
iterations (9600 models) to sample the posterior PDF,  $P(\mmin,
\mmax, f_{\rm occ}| \mathcal{M}) \propto \exp(-\chi_{\mathcal
  M}^2/2)$, with

@@

In Fig. \ref{fig:like} we present yhe one and two dimensional
projections of the posterior probability distributions of \mmin, \mmax
and \focc. We find that the \mmin and \mmax parameter space cannot be
tightly constrained mainly because of cosmic variance and the large
observational Poisonnian uncertainty in the ACF. We however found that
9.6$<\log \mmin<$11.0 and 10.9$<\log\mmax<$13.0.   We interestingly
find that \focc\ is basically completely determined by \mmin  going
from \focc$=$0.015 when $\log\mmin=$9.6 to \focc$=$0.30 when
$\log\mmin=$11.0
($\focc\sim0.1\left(\mmin/10^{10.5}\right)^{0.93}$). We remark that
those models where $\log \focc>0.00$ ($\focc>1$) correspond  to models where the number density  of halos is smaller than the number of observed LAEs but are
still consistent because of the uncertainty in the median number
density of LAEs in the universe due to cosmic variance (see
Fig. \ref{fig:cosmicv0}).%in the number of dark matter halos along the
                         %mock fields

@@

 In Fig. \ref{fig:mmed} the green vertical lines represent  the 16, 50
 and 84 percentiles  of the posterior probability distribution of
 \mmin that we previously obtained. From this figure we find that the
 median dark matter halo mass is $\tilde{M}_{\mathrm h} =
 10^{10.6^{+0.7}_{-0.6}} $. 
This result is consistent with previous
 estimations of  the  median dark matter halo mass reported by
 \citet{Bielby16} ($\tilde{M}_{\mathrm h} = 10^{11.0\pm0.6}$),
 \citet{Gawiser07} ($\tilde{M}_{\mathrm h} = 10^{10.9\pm0.9}$) and
 \citet{Ouchi2010} ($\tilde{M}_{\mathrm h} =
 6.7^{+42.0}_{6.7}\times10^{10}$) using semi-analitical approaches.

@@

\begin{figure}
\includegraphics[width=0.46\textwidth, height=0.33\textwidth]{mmin_dfocc1.png}
\caption{Width of the halo number distribution over the 54 mock field of the simulation ($\Delta \log \left( N_{\mathrm halos}/\tilde{N}_{\mathrm halos}\right)$) of the central 68 (blue diamonds) and  95(red circles)  percentiles vs $M_{\rm min}$. Points are color coded according to their associated value of $\Delta M \equiv M_{\rm min}-M_{\rm min}$. The green line diagonal line accounts for the 1:1 relation and the green vertical lines represent  the 16, 50 and 84 percentiles  of the posterior probability distribution of \mmin.}
\label{fig:cosmicv}
\end{figure}

@@

In this letter we studied the impact of cosmic variance and observational uncertainties in constraining the mass range and  occupation fraction of dark matter halos hosting  LAEs.  
To this end, we used the BolshoiP N-body simulation to construct  27 mock fields  with the same typical size  of observed fields at  $z=3.1$ ($\sim 1 {\rm deg^2}$).  
In our model a dark matter halo with mass in the range $M_{\rm min}<M_{\mathrm h}<M_{\rm   max}$ can only host one detectable LAE at most. 
We explored the parameter space determined by \mmin\ and \mmax\ using Monte Carlo Markov-Chain minimization to match the observed  ACF and mean number density of LAEs.
